---
title: "Word Relationships"
subtitle: "Investigating the Harry Potter Book Series"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  html_document:
    df_print: 'paged'
    css: '../css/logm655.css'
    toc: yes
    toc_float: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = F,
                      warning = F,
                      fig.align = 'center',
                      cache = T)
```

# Overview

The term text mining (aka text analysis) describes a process wherein relevent and/or interesting information is extracted from a corpus of documents. Text is often in an unstructured format so performing even the most basic analysis requires some re-structuring.  Thus, an important question to ask before starting any text analysis project is "To what format should I convert my text data to best support the analyses I want to perform?"  In R, there are two primay format to choose from: the tidy data format using the `tidytext` package and the non-tidy data format supported by several packages including `tm`, `text2vec`, `quanteda`, and `RWeka`.

The `tidytext` package allows us to analyze the text using the __*tidy text*__ format: a table with one-token-per-document-per-row. This format allows us to efficiently pipe our analysis directly into the popular suite of `tidyverse` tools such as `dplyr`, `tidyr`, and `ggplot2` to explore and visualize the data. Most text analysis/NLP tools in R do not use a tidy text format. The [CRAN Task View for Natural Language Processing](https://cran.r-project.org/web/views/NaturalLanguageProcessing.html) lists a large selection of packages that take other structures of input and provide non-tidy outputs. These packages are very useful in text mining applications, and many existing text datasets are structured according to these formats. Thus, its extremely important to understand how to convert back-and-forth between different formats.

This document presents several methods analyze text data and leverages the data provided in the [`harrypotter` package](https://github.com/bradleyboehmke/harrypotter) created by Brad Boehmke. This package has not been published to the CRAN, but can be installed from GitHub by running the code below:

```{r, eval=FALSE}
if (packageVersion("devtools") < 1.6) {
  install.packages("devtools")
}

devtools::install_github("bradleyboehmke/harrypotter")
```   

This package provides the text from the following novels in the Harry Potter series:

- `philosophers_stone`: Harry Potter and the Philosophers Stone (1997)
- `chamber_of_secrets`: Harry Potter and the Chamber of Secrets (1998)
- `prisoner_of_azkaban`: Harry Potter and the Prisoner of Azkaban (1999)
- `goblet_of_fire`: Harry Potter and the Goblet of Fire (2000)
- `order_of_the_phoenix`: Harry Potter and the Order of the Phoenix (2003)
- `half_blood_prince`: Harry Potter and the Half-Blood Prince (2005)
- `deathly_hallows`: Harry Potter and the Deathly Hallows (2007)
 to illustrate various text mining and analysis capabilities.  

The text from each book is stored as a character vector with each element representing a single chapter.  For instance, the following illustrates the raw text of the first chapter of the `philosophers_stone`:

```{r}
harrypotter::philosophers_stone[1]
```

Additionally, we'll need to install and load the following packages to help with this analysis.

```{r, message=FALSE, warning = F}
pacman::p_load(tm, 
               pdftools, 
               here,
               tau,
               tidyverse,
               stringr,
               tidytext, 
               RColorBrewer,
               qdap,
               qdapRegex,
               qdapDictionaries,
               qdapTools,
               data.table,
               coreNLP,
               scales,
               harrypotter,
               text2vec,
               SnowballC,
               DT,
               quanteda,
               RWeka,
               broom,
               tokenizers,
               grid,
               knitr,
               widyr)

pacman::p_load_gh("dgrtwo/drlib",
                  "trinker/termco", 
                  "trinker/coreNLPsetup",        
                  "trinker/tagger")
```

```{r child_tidy, child='text_mining_tidy.Rmd'}
```

```{r child_idf, child='text_mining_tf_idf.Rmd', eval=!FALSE}
```

```{r child_relate, child='text_mining_word_relationships.Rmd', eval=!FALSE}
```