---
title: "OPER 655 Student Project Report"
author: "Your name"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  html_document:
    code_folding: 'hide'
abstract: 'This is where you put your abstract'
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = F,
                      warning = F, 
                      comment = NA)
```

## Project Overview/Problem Statement 

Describe your project in sufficient detail for someone who is not familiar with text mining or NLP can understand what you are saying. This section should conclude with a problem statement that clearly and concisely describes the problem you are addressing (aka the question you are trying to answer). 

### Installing Loading Required R Packages

Before going too far, be sure your reader has the necessary packages to follow along with your methodology and ultimately reproduce your results.

```{r, warning=FALSE, message=FALSE}
pacman::p_load(tm,
               tidytext)
```

## Methodology 

Describe the steps and methods you used for your analysis to go from raw data to end result.

### Data 

Describe the data

- Where did you get the data?
- In what format was the data when you received/retreived it?
- What transformations did you perform on the data?

### Analysis

Walk your reader through you analysis.  The yaml header is already set to fold all code chunks. 

## Findings and Conclusions 

Describe your findings and conclusions here.  Include your important visualizations

### Future Work

What else would you do if you had more time?

