---
title: "OPER 655 Student Project Report"
author: "Tyler Spangler"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: 
  html_document:
    code_folding: 'hide'
bibliography: bib.bib
abstract: The purpose of this report is to analyze the last statments from Texas Prisoner inmates.  The data set from @kaggle contains data on 545 executions with prisoner attributes including race, education level, age, etc.  This research will utilize topic modeling and sentiment analysis to determine if there are similarities among last statements between prisoners.   

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = F,
                      warning = F, 
                      comment = NA)
```

## Project Overview/Problem Statement 

Outline:
1.    

Describe your project in sufficient detail for someone who is not familiar with text mining or NLP can understand what you are saying. This section should conclude with a problem statement that clearly and concisely describes the problem you are addressing (aka the question you are trying to answer). 

### Installing Loading Required R Packages

Before going too far, be sure your reader has the necessary packages to follow along with your methodology and ultimately reproduce your results.

```{r, warning=FALSE, message=FALSE}
pacman::p_load(tm, 
               pdftools, 
               here,
               tau,
               tidyverse,
               stringr,
               tidytext, 
               RColorBrewer,
               qdap,
               qdapRegex,
               qdapDictionaries,
               qdapTools,
               data.table,
               coreNLP,
               scales,
               harrypotter,
               text2vec,
               SnowballC,
               DT,
               quanteda,
               RWeka,
               broom,
               tokenizers,
               grid,
               knitr,
               widyr,
               textdata,
               tidyr,
               topicmodels)
```

## Methodology 

Describe the steps and methods you used for your analysis to go from raw data to end result.

### Data 

Describe the data

- Where did you get the data?
- In what format was the data when you received/retreived it?
- What transformations did you perform on the data?

### Analysis

Walk your reader through you analysis.  The yaml header is already set to fold all code chunks. 

## Findings and Conclusions 

Describe your findings and conclusions here.  Include your important visualizations

### Future Work

What else would you do if you had more time?

